The Construction of Set-Truncated Higher Inductive Types

### 2019-03-31 paper/paper.tex
Futher work on paper

% Category theory
\newcommand{\idm}[1]{\morphism{id}_{#1}} % identity morphism
\newcommand{\compm}[2]{#2 \circ #1} % composition of morphisms
\newcommand{\iso}[2]{#1 \cong #2} % isomorphisms
\newcommand{\setoids}{\category{Setoid}} % category of setoids
\newcommand{\functor}[2]{#1 \longrightarrow #2} % functors
\newcommand{\idf}[1]{\functortxt{id}_{#1}} % identity functor
\newcommand{\Cf}[1]{\functortxt{C}_{#1}} % constant functor
\newcommand{\compf}[2]{#2 \circ #1} % composition of functors
\newcommand{\prodf}[2]{#1 \times #2} % product of functors
\newcommand{\sumf}[2]{#1 + #2} % sum of functors
\newcommand{\nattrans}[2]{#1 \Longrightarrow #2} % natural transfomations
\newcommand{\ctrans}[1]{\nattranstxt{c}_{#1}} % constant transformation
\newcommand{\constrt}[1]{\nattranstxt{constr}_{#1}} % constructor transformation
\newcommand{\idt}[1]{\nattranstxt{id}} % identity transformation
\newcommand{\compt}[2]{#2 \circ #1} % composition of natural transformations
\newcommand{\lwhisker}[2]{#2 \vartriangleright #1} % whiskering of transformations
\newcommand{\inlt}{\nattranstxt{inl}} %left inclusion of transformations
\newcommand{\inrt}{\nattranstxt{inr}} % right inclusion of transformations
\newcommand{\prlt}{\nattranstxt{pr}_1} % first projection of transformations
\newcommand{\prrt}{\nattranstxt{pr}_2} % second projection of transformations
\newcommand{\pairt}[2]{(#1 , #2)} % pairing of transformations
\newcommand{\idtoiso}{\function{idtoiso}} % identity to iso
\newcommand{\falg}[1]{\category{FAlg}(#1)} % category of algebras on a functor

@@

% Examples of HITs
\newcommand{\Circ}{S^1} % circle
\newcommand{\base}{\constructor{base}} % base point of circle
\newcommand{\loopS}{\constructor{loop}} % base point of circle
\newcommand{\modZ}{\mathbb{Z}_2} % integers modulo 2
\newcommand{\ZZ}{\constructor{Z}} % zero
\newcommand{\SZ}{\constructor{S}} % zero
\newcommand{\modP}{\constructor{mod}} % n = n + 2
\newcommand{\HIT}{\function{HIT}} % the HIT of a signature

\newcommand{\remove}{}

\def\lastname{Van der Weide}
\begin{document}
\begin{frontmatter}
  \title{The Construction of Set-Truncated Higher Inductive Types} \author{Niels van der Weide
    \thanksref{ALL}\thanksref{myemail}}
  \address{Institute for Computation and Information Sciences\\ Radboud Universiteit\\
    Nijmegen, The Netherlands} \author{My Co-author\thanksref{coemail}}
  \address{My Co-author's Department\\My Co-author's University\\
    My Co-author's City, My Co-author's Country} \thanks[ALL]{We would like to thank Dan Frumin and Andrej Bauer for inspiring discussions. Work on this article was supported by a grant from the \href{https://eutypes.cs.ru.nl/}{COST Action EUTypes CA15123}.} \thanks[myemail]{Email:
    \href{mailto:nweide@cs.ru.nl} {\texttt{\normalshape
        nweide@cs.ru.nl}}} \thanks[coemail]{Email:
    \href{mailto:couserid@codept.coinst.coedu} {\texttt{\normalshape
        couserid@codept.coinst.coedu}}}
\begin{abstract} 
  We construct finitary set-truncated higher inductive types (HITs) using quotients and the propositional truncation.
  Using a modification of the schema by Basold \etal, we define a notion of signatures and univalent categories of algebras in both sets and setoids.
  To interpret HITs, we use initial algebra semantics for which we show that the initiality implies induction.
  The desired algebra is obtained by lifting the quotient adjunction to the level of algebras and adapting Dybjer's and Moenclaey's interpretation of HITs in setoids.
  From this construction, we conclude that the equality types of HITs are freely generated and that HITs are unique.
  The results are formalized in the UniMath library.
\end{abstract}
\begin{keyword}
higher inductive types, homotopy type theory, category theory, setoids, intuitionistic type theory, Coq
\end{keyword}
\end{frontmatter}
\section{Introduction}
\label{intro}
Homotopy type theory (HoTT) is a form of intensional type theory. 
It has semantics in the simplicial sets model \cite{simpset}, and types represent spaces, terms represent points, and equalities represent paths.
Furthermore, equality is proof relevant and we frequently talk about homotopies: paths between paths.

One of the main features of HoTT, is higher inductive types (HITs) \cite{hottbook}.
These are types generated by constructors for their points, paths, homotopies, and so on.
HITs have been used in numerous applications among which are homotopical patch theory \cite{AngiuliMLH16}. synthetic homotopy theory \cite{LicataF14,LicataS13,hottbook}, defining type theory within type theory \cite{AltenkirchK16}, constructive finiteness \cite{FruminGGW18}, and the partiality monad \cite{AltenkirchDK17}.

To get a feeling for what HITs are, let us look a some examples before looking at them more formally.

@@

The first one, $\Circ$, is the \emph{circle}.
It has a point, $\base$, and a path $\loopS : \base = \base$.
Since neither the point nor path constructors uses arguments from $\Circ$, this HIT is \emph{nonrecursive}.
The second, $\trunct{A}$, is the \emph{propositional truncation} of $A$.
This type is $A$ with all its points identified.
Note that the path constructor $p$ uses arguments from $\trunct{A}$, and thus this HIT is called \emph{recursive}.
The last one $, \modZ$ is the \emph{integers modulo 2} and note that both the point constructor $\SZ$ and path constructor $\modP$ are recursive.

HITs have also been studied more formally and several schemes have already been developed.
Awodey \etal \ showed that the rules of inductive types \cite{dybjer1994inductive} in HoTT are equivalent to being a homotopy initial algebra \cite{AwodeyGS12}.
They extended this result to a class of HITs called \emph{W-suspensions} \cite{Sojakova15}.
Another scheme, quotient inductive-inductive types, was defined semantically by Altenkirch \etal \ and they prove initial algebra semantics \cite{AltenkirchCDKF18}.
Assuming UIP, Kaposi \etal \ constructed finitary QIITs as the syntax of a domain specific type theory \cite{KaposiKA19}, and they extended this scheme to higher inductive-inductive types \cite{KaposiK18}.
Basold \etal \ defined an external syntactic scheme for HITs with only point and path constructors \cite{BasoldGW17}.
Dybjer and Moenclaey defined a scheme allowing constructors for points, paths, and homotopies. and give its semantics in the groupoid model  \cite{DybjerM18,HofmannS94}.
Cavallo and Harper extend computational cubical type theory with a scheme for indexed cubical inductive types and proved canonicity \cite{AngiuliHW17,CavalloH19}.

Furthermore, the semantics of HITs also has been studied.
Coquand \etal \ provide the semantics of various higher inductive types, such as spheres, the torus, the truncation, and the pushout, in cubical type theory \cite{BezemCH13,CohenCHM16,CoquandHM18}.
On the other hand, Lumsdaine and Shulman also study the semantics of higher inductive types, but they use a semantic scheme (cell monads with parameters) and proof existence in sufficiently nice Quillen model categories \cite{lumsdaine2017semantics}.
Some recursive higher inductive types can be constructed from nonrecursive HITs.
Both Kraus and Van Doorn show that the propositional truncation can be constructed from non-recursive HITs \cite{Kraus16,Doorn16} and Rijke shows that every truncation can be constructed via non-recursive HITs \cite{rijke2017join}.

In this paper, we show that all finitary set-truncated HITs can be constructed with quotients and propositional truncations. 
This briefly means that all higher inductive types exist if a small number of simple ones exist.
Furthermore, we use this construction to deduce several properties of HITs such as uniqueness and that the path space is freely generated.
The latter result is similar to work by Kraus and Von Raumer \cite{KrausRaumer}.
Note that from this construction we can conclude that finitary set-truncated HITs can be constructed from non-recursive HITs.
Since quotients can be constructed from coequalizers and set truncations, we only need coequalizers and propositional/set truncations for this construction.

The main idea of the proof is to take advantage that the quotient is an adjoint functor \cite{RijkeS15} from setoids to sets.
To relate this to HITs, we first define schemes, which are an internalized version of the schema by Basold \etal \ \cite{BasoldGW17}, and categories of algebras on them.
Then we lift the quotient adjunction to an adjunction from algebras in setoids and algebras in sets.
Since initiality implies induction, it suffices to construct the initial setoid algebra for which we adapt the construction by Dybjer and Moenclaey \cite{moeneclaey2016schema,DybjerM18}.

@@

\section{Constructing the Initial Algebra}
\label{sec:construction}

\subsection{Adjunction between Algebras}
To construct the adjunction between the category of algebras, we use two lemmata

\begin{lemma}
Given are categories $\mathcal{C}$ and $\mathcal{D}$, functors $A_{\mathcal{C}} : \functor{\mathcal{C}}{\mathcal{C}}$, $A_{\mathcal{D}} : \functor{\mathcal{D}}{\mathcal{D}}$, and $F : \functor{\mathcal{C}}{\mathcal{D}}$.
Then we get a functor 
\end{lemma}

\begin{lemma}
Given are categories $\mathcal{C}$ and $\mathcal{D}$, functors $A_{\mathcal{C}} : \functor{\mathcal{C}}{\mathcal{C}}$, $A_{\mathcal{D}} : \functor{\mathcal{D}}{\mathcal{D}}$, and an adjunction ....
Suppose, the following diagrams commute

Then we get an adjunction
\end{lemma}

\begin{lemma}
Adjunction between full subcategories
\end{lemma}

Now we apply these to get an adjunction between set HIT-algebras and setoid HIT-algebras.
For this, we do the following steps.

\begin{lemma}
The path setoid commutes with sums and products.

The quotient commutes with sums and products.
\end{lemma}

\begin{lemma}
We have a natural isomorphism (path setoid/quotients commutes with polynomials).
\end{lemma}

@@

\begin{lemma}
There is an adjunction between prealgebras ...
\end{lemma}

@@

All in all, we get

\begin{proposition}
We have an adjunction
\end{proposition}

\subsection{Initial Setoid Algebra}

Hence, we conclude

\begin{theorem}
\label{thm:hit-exist}
HITs exist
\end{theorem}

@@

\begin{corollary}
If $H_1$ and $H_2$ are HITs for $\sig$, then the underlying $\sig$-algebras of $H_1$ and $H_2$ are equal.
\end{corollary}

@@

\begin{corollary}
Let $\sig$ be a signature.
Then for each $x, y : ...$ we have an equivalence $....$.
\end{corollary}

\section{Conclusion and Further Work}
\label{sec:conclusion}
\cite{ahrens2019bicategories}
\cite{capriotti2017univalent}

@@

	\bibitem{Doorn16}
	Floris van Doorn, \emph{{Constructing the Propositional Truncation using
			Non-Recursive HITs}}, Proceedings of the 5th {ACM} {SIGPLAN} Conference on
	Certified Programs and Proofs, Saint Petersburg, FL, USA, January 20-22,
	2016, 2016, pp.~122--129.	
\end{thebibliography}

%\begin{thebibliography}{10}\label{bibliography}
%\bibitem{cy} Civin, P., and B. Yood, \emph{Involutions on Banach
%    algebras}, Pacific J. Math. \textbf{9} (1959), 415--436.

%\bibitem{cp} Clifford, A. H., and G. B. Preston, ``The Algebraic
%  Theory of Semigroups,'' Math. Surveys \textbf{7}, Amer. Math. Soc.,
%  Providence, R.I., 1961.

%\bibitem{f} Freyd, Peter, Peter O'Hearn, John Power, Robert Tennent
%  and Makoto Takeyama, \emph{Bireflectivity}, Electronic Notes in
%  Theoretical Computer Science {\bf 1} (1995), URL:
%  \href{https://www.sciencedirect.com/journal/electronic-notes-in-theoretical-computer-science/vol/1/suppl/C}
%  {\texttt{http://www.elsevier.com/locate/entcs/volume1.html}}.

%\bibitem{em2} Easdown, D., and W. D. Munn, \emph{Trace functions on
%    inverse semigroup algebras}, U. of Glasgow, Dept. of Math.,
%  preprint 93/52.

%\bibitem{r} Roscoe, A. W., ``The Theory and Practice of Concurrency,''
%  Prentice Hall Series in Computer Science, Prentice Hall Publishers,
%  London, New York (1198), 565pp. With associated web site\\  
%  \href{http://www.comlab.ox.ac.uk/oucl/publications/books/concurrency/}
%  {\texttt{http://www.comlab.ox.ac.uk/oucl/publications/books/concurrency/}}.

%\bibitem{s} Shehadah, A. A., ``Embedding theorems for semigroups with
%  involution, `` Ph.D.  thesis, Purdue University, Indiana, 1982.

%\bibitem{w} Weyl, H., ``The Classical Groups,'' 2nd Ed., Princeton U.
%  Press, Princeton, N.J., 1946.

%\end{thebibliography}
\end{document}

>>>>

% Category theory
\newcommand{\idm}[1]{\morphism{id}_{#1}} % identity morphism
\newcommand{\compm}[2]{#2 \circ #1} % composition of morphisms
\newcommand{\iso}[2]{#1 \cong #2} % isomorphisms
\newcommand{\setoids}{\category{Setoid}} % category of setoids
\newcommand{\functor}[2]{#1 \longrightarrow #2} % functors
\newcommand{\idf}[1]{\functortxt{id}_{#1}} % identity functor
\newcommand{\Cf}[1]{\functortxt{C}_{#1}} % constant functor
\newcommand{\compf}[2]{#2 \circ #1} % composition of functors
\newcommand{\prodf}[2]{#1 \times #2} % product of functors
\newcommand{\sumf}[2]{#1 + #2} % sum of functors
\newcommand{\nattrans}[2]{#1 \Longrightarrow #2} % natural transfomations
\newcommand{\ctrans}[1]{\nattranstxt{c}_{#1}} % constant transformation
\newcommand{\constrt}[1]{\nattranstxt{constr}_{#1}} % constructor transformation
\newcommand{\idt}[1]{\nattranstxt{id}} % identity transformation
\newcommand{\compt}[2]{#2 \circ #1} % composition of natural transformations
\newcommand{\lwhisker}[2]{#2 \vartriangleright #1} % whiskering of transformations
\newcommand{\inlt}{\nattranstxt{inl}} %left inclusion of transformations
\newcommand{\inrt}{\nattranstxt{inr}} % right inclusion of transformations
\newcommand{\prlt}{\nattranstxt{pr}_1} % first projection of transformations
\newcommand{\prrt}{\nattranstxt{pr}_2} % second projection of transformations
\newcommand{\pairt}[2]{(#1 , #2)} % pairing of transformations
\newcommand{\idtoiso}{\function{idtoiso}} % identity to iso
\newcommand{\falg}[1]{\category{FAlg}(#1)} % category of algebras on a functor
\newcommand{\adj}[2]{#1 \dashv #2} % adjunction
\newcommand{\fullsub}[2]{\function{full}(#1,#2)} % full subcategory
\newcommand{\factorsub}[1]{#1_{\function{sub}}} % factor through full sub

@@

% Examples of HITs
\newcommand{\Circ}{S^1} % circle
\newcommand{\base}{\constructor{base}} % base point of circle
\newcommand{\loopS}{\constructor{loop}} % base point of circle
\newcommand{\modZ}{\mathbb{Z}_2} % integers modulo 2
\newcommand{\ZZ}{\constructor{Z}} % zero
\newcommand{\SZ}{\constructor{S}} % successor
\newcommand{\modP}{\constructor{mod}} % n = n + 2
\newcommand{\HIT}{\function{HIT}} % the HIT of a signature

% Adjunction
\newcommand{\quotF}{\mathcal{Q}}
\newcommand{\pathF}{\pi_0}
\newcommand{\quotPA}{\mathcal{Q}^{\category{PreAlg}}}
\newcommand{\pathPA}{\pi^{\category{PreAlg}_0}}
\newcommand{\quotAA}{\mathcal{Q}^{\category{Alg}}}
\newcommand{\pathAA}{\pi_0^{\category{Alg}}}

% Free congruence
\newcommand{\CRC}{\mathcal{X}} % carrier
\newcommand{\CR}{\mathcal{R}^*} % relation
\newcommand{\CRR}{\mathcal{R}} % relation
\newcommand{\CRrefl}{\constructor{refl}} % reflexivity
\newcommand{\CRsym}{\constructor{sym}} % symmetry
\newcommand{\CRtrans}{\constructor{trans}} % transitivity
\newcommand{\CRpair}{\constructor{pair}} % pair
\newcommand{\CRinl}{\constructor{inl}} % in left
\newcommand{\CRinr}{\constructor{inr}} % in right
\newcommand{\CRpath}{\constructor{path}} % path constuctors
\newcommand{\CRcong}{\constructor{cong}} % congruence

\newcommand{\remove}{} % remove a part

\def\lastname{Van der Weide}
\begin{document}
\begin{frontmatter}
  \title{The Construction of Set-Truncated Higher Inductive Types} \author{Niels van der Weide
    \thanksref{ALL}\thanksref{myemail}}
  \address{Institute for Computation and Information Sciences\\ Radboud Universiteit\\
    Nijmegen, The Netherlands} \author{My Co-author\thanksref{coemail}}
  \address{My Co-author's Department\\My Co-author's University\\
    My Co-author's City, My Co-author's Country} \thanks[ALL]{We would like to thank Dan Frumin and Andrej Bauer for inspiring discussions. Work on this article was supported by a grant from the \href{https://eutypes.cs.ru.nl/}{COST Action EUTypes CA15123}.} \thanks[myemail]{Email:
    \href{mailto:nweide@cs.ru.nl} {\texttt{\normalshape
        nweide@cs.ru.nl}}} \thanks[coemail]{Email:
    \href{mailto:couserid@codept.coinst.coedu} {\texttt{\normalshape
        couserid@codept.coinst.coedu}}}
\begin{abstract} 
  We construct finitary set-truncated higher inductive types (HITs) using quotients and the propositional truncation.
  Using a modification of the schema by Basold \etal, we define a notion of signatures and univalent categories of algebras in both sets and setoids.
  To interpret HITs, we use initial algebra semantics for which we show that the initiality implies induction.
  The desired algebra is obtained by lifting the quotient adjunction to the level of algebras and adapting Dybjer's and Moeneclaey's interpretation of HITs in setoids.
  From this construction, we conclude that the equality types of HITs are freely generated and that HITs are unique.
  The results are formalized in the UniMath library.
\end{abstract}
\begin{keyword}
higher inductive types, homotopy type theory, category theory, setoids, intuitionistic type theory, Coq
\end{keyword}
\end{frontmatter}
\section{Introduction}
\label{intro}
Homotopy type theory (HoTT) is a form of intensional type theory. 
It has semantics in the simplicial sets model \cite{simpset}, and types represent spaces, terms represent points, and equalities represent paths.
Furthermore, equality is proof relevant and we frequently talk about homotopies: paths between paths.

One of the main features of HoTT, is higher inductive types (HITs) \cite{hottbook}.
These are types generated by constructors for their points, paths, homotopies, and so on.
HITs have been used in numerous applications among which are homotopical patch theory \cite{AngiuliMLH16}. synthetic homotopy theory \cite{LicataF14,LicataS13,hottbook}, defining type theory within type theory \cite{AltenkirchK16}, constructive finiteness \cite{FruminGGW18}, and the partiality monad \cite{AltenkirchDK17}.
To get a feeling for what HITs are, let us look a some examples.

@@

The first one, $\Circ$, is the \emph{circle}.
It has a point $\base$ and a path $\loopS : \base = \base$.
Since neither the point nor path constructors uses arguments from $\Circ$, this HIT is \emph{nonrecursive}.
The second, $\trunct{A}$, is the \emph{propositional truncation} of $A$.
This type is $A$ with all its points identified.
Note that the path constructor $p$ uses arguments from $\trunct{A}$, which means this HIT is \emph{recursive}.
The last one, $\modZ$, is the \emph{integers modulo 2} and note that both the point constructor $\SZ$ and path constructor $\modP$ are recursive.

HITs have been studied formally and several schemes have already been developed.
Awodey \etal \ showed that the rules of inductive types \cite{dybjer1994inductive} in HoTT are equivalent to being a homotopy initial algebra \cite{AwodeyGS12}.
They extended this result to a class of HITs called \emph{W-suspensions} \cite{Sojakova15}.
Another scheme, quotient inductive-inductive types (QIITs), was defined semantically by Altenkirch \etal \ and they show initial algebras are QIITs \cite{AltenkirchCDKF18}.
Assuming UIP, Kaposi \etal \ constructed finitary QIITs as the syntax of a domain specific type theory \cite{KaposiKA19}, and they extended this scheme to higher inductive-inductive types \cite{KaposiK18}.
Basold \etal \ defined an external syntactic scheme for HITs with only point and path constructors \cite{BasoldGW17}
while Dybjer and Moenclaey defined a scheme allowing constructors for points, paths, and homotopies, and they give its interpretation in the groupoid model  \cite{DybjerM18,HofmannS94}.
Cavallo and Harper extend computational cubical type theory with a scheme for indexed cubical inductive types and proved canonicity \cite{AngiuliHW17,CavalloH19}.

Furthermore, the semantics of HITs also has been studied.
Coquand \etal \ provide the semantics of various higher inductive types, such as spheres, the torus, the truncation, and the pushout, in cubical type theory \cite{BezemCH13,CohenCHM16,CoquandHM18}.
On the other hand, Lumsdaine and Shulman also study the semantics of higher inductive types using a semantic scheme (cell monads with parameters) and they prove existence in sufficiently nice Quillen model categories \cite{lumsdaine2017semantics}.
Some recursive higher inductive types can be constructed from nonrecursive HITs.
Both Kraus and Van Doorn show that the propositional truncation can be constructed from non-recursive HITs \cite{Kraus16,Doorn16} and Rijke shows that every truncation can be constructed via non-recursive HITs \cite{rijke2017join}.

In this paper, we show that all finitary set-truncated HITs can be constructed with quotients and propositional truncations. 
This briefly means that all higher inductive types exist if a small number of simple ones exist.
A similar result was obtained by Kaposi \etal \ \cite{KaposiKA19}, but we do not assume UIP and we use simpler HITs for the construction.
Furthermore, we use this construction to deduce several properties of HITs such as uniqueness and that the path space is freely generated.
Kraus and Von Raumer obtained an analogous result for the coequalizer by providing a nicer induction principle for its path types \cite{KrausRaumer}.
Note that from our construction we can conclude that all finitary set-truncated HITs can be constructed from non-recursive HITs.
Since quotients can be constructed from coequalizers and set truncations, we only need coequalizers and propositional/set truncations for this construction \cite{rijke2017join}.

The main idea of the proof is to take advantage that the quotient is left adjoint functor from setoids to sets \cite{RijkeS15}.
To relate this to HITs, we first define schemes, which are an internalized version of the schema by Basold \etal \ \cite{BasoldGW17}, and categories of algebras on them.
Then we lift the quotient adjunction to an adjunction from algebras in setoids and algebras in sets.
Since initiality implies induction, it suffices to construct the initial setoid algebra for which we adapt the construction by Dybjer and Moenclaey \cite{moeneclaey2016schema,DybjerM18}.

@@

\section{Constructing the Initial Algebra}
\label{sec:construction}
To construct the initial algebra, we first define a left adjoint from algebras in setoids to algebras in algebras.
Such functors preserve initial objects, and thus it suffices to construct the initial setoid algebra.
For that, we use an adaption of Dybjer's and Moeneclaey's interpretation of HITs in the setoid model \cite{DybjerM18,moeneclaey2016schema}.
In the remainder of this section, we work with a fixed signature $\sig$.

\subsection{Quotient Adjunction}
We start by defining a left adjoint functor $\quotF : \functor{\setoids}{\hset}$.
On objects, we define $\quotF \> (X , R) = \setquot{X}{R}$ and its action on morphisms is defined by recursion on the quotient type.
The laws are proven by quotient induction.
The right adjoint is the path setoid functor $\pathF : \functor{\hset}{\setoids}$.
For sets $X$, we define functor $\pathF \> X$ to be $(X, =_X)$ where $=_X$ is the equivalence relation on $X$ defined by $\lambda x y, x = y$.
Note that this action is functorial and that this gives rise to an adjunction.

\begin{lemma}[Theorem 2.20 from \cite{RijkeS15}]
We have an adjunction $\adj{\quotF}{\pathF}$.
\end{lemma}

\subsection{The Adjunction on Prealgebras}
With this adjunction in place, our first step is to lift it to the level of prealgebras.
The first step, is to lift the functors for which we use the following lemma.

\begin{lemma}
\label{lem:prealgadj}
Given are categories $\mathcal{C}$ and $\mathcal{D}$, functors $A_1 : \functor{\mathcal{C}}{\mathcal{C}}$, $A_2 : \functor{\mathcal{D}}{\mathcal{D}}$, and $F : \functor{\mathcal{C}}{\mathcal{D}}$, and a natural transformation $n : \nattrans{\compf{F}{A_2}}{\compf{A_1}{F}}$.
Then we get a functor $F^{\category{PreAlg}} : \functor{\falg{A_1}}{\falg{A_2}}$.
\end{lemma}

For algebras $(X, f) : \falg{A_1}$, we define $F^{\category{PreAlg}}(X,f) = (F \> X, \compm{F \> f}{n \> X})$.

\begin{lemma}
\label{lem:comm}
We have
\begin{itemize}
	\item $\pathF$ commutes with sums and products.
	\item $\quotF$ commutes with sums and products.
\end{itemize}
This gives rise to two natural isomorphisms $n_1 : \nattrans{\compf{\pathF}{\semPT{\pointc{\sig}}}}{\compf{\semP{\pointc{\sig}}}{\pathF}}$
\end{lemma}

Now we use Lemmata \ref{lem:prealgadj} and \ref{lem:comm} to lift both the quotient and the path setoid functors and we get
\[
\quotPA : \functor{\prealgst{\sig}}{\prealg{\sig}}
\quad \quad
\pathPA : \functor{\prealg{\sig}}{\prealgst{\sig}}.
\]

\begin{lemma}
We have an adjunction $\adj{\quotPA}{\pathPA}$.
\end{lemma}

\subsection{Algebras}

If $\mathcal{C}$ is a category and $P$ is a family of propositions on $\mathcal{C}$, then we write $\fullsub{\mathcal{C}}{P}$ for the full subcategory of $\mathcal{C}$ in which each object satisfies $P$.
First, we look at way to obtain adjunctions between full subcategories.

\begin{lemma}
Let $\mathcal{C}$ and $\mathcal{D}$ be categories, let $P_1$ and $P_2$ be families of propositions on $\mathcal{C}$ and $\mathcal{D}$ respectively, and suppose we have a functor $F : \functor{\mathcal{C}}{\mathcal{D}}$.
If for each object $x : \mathcal{C}$ we have $P_1 \> x \rightarrow P_2(F \> x)$, then we get a functor $\factorsub{F} : \functor{\fullsub{\mathcal{C}}{P_1}}{\fullsub{\mathcal{D}}{P_2}}$.
\end{lemma}

\begin{proposition}
Let $\mathcal{C}$ and $\mathcal{D}$ be categories, let $P_1$ and $P_2$ be families of propositions on $\mathcal{C}$ and $\mathcal{D}$ respectively, and suppose we have an adjunction $\adj{L}{R}$ with $L : \functor{\mathcal{C}}{\mathcal{D}}$.
If for each object $x : \mathcal{C}$ we have $P_1 \> x \rightarrow P_2(L \> x)$ and for each $y : \mathcal{D}$ we have $P_2 \> x \rightarrow P_1(R \> x)$, then we get an adjunction $\adj{\factorsub{L}}{\factorsub{R}}$.
\end{proposition}

Now we would like to apply this proposition to the adjunction obtained from Lemma \ref{lem:prealgadj}.

\begin{theorem}
\label{thm:algadj}
We have an adjunction $\adj{\quotAA}{\pathAA}$.
\end{theorem}

@@

@@

\subsection{Initial Setoid Algebra}
The initial setoid is constructed in two steps.
First we define its carrier as the initial algebra on $\pointc{\sig}$ and then we define the equivalence relation as the least congruence relation containing the equations in $\sig$ and preserving equality of the point constructor.

\begin{lemma}
The category $\prealg{\pointc{\sig}}$ has an initial object.
\end{lemma}

This follows from Ad\'amek's theorem and the fact that the functors $\semP{P}$ is $\omega$-continuous for each $P$.
We denote the initial object of this category by $(\CRC , f)$.
The carrier of the desired setoid is $\CRC$ meaning that the next step is to define the equivalence relation on $\CRC$.

The main difficulty of defining this relation, is that we also must behave correctly on $\semP{\pointc{\sig}}{\CRC}$, because otherwise, we would not be able to lift $f$ to a setoid morphism.
For that reason, we first define a relation $\CR \> P$ on $\semP{P} \> \CRC$ for each $P : \poly$..

\begin{center}
\begin{bprooftree}
\AxiomC{$x : \semP{P} \> \CRC$}
\UnaryInfC{$\CRrefl \> x : \CR \> P \> x \> x$}
\end{bprooftree}
\begin{bprooftree}
\AxiomC{$r : \CR \> P \> x \> y$}
\UnaryInfC{$\CRsym \> r : \CR \> P \> y \> x$}
\end{bprooftree}
\begin{bprooftree}
\AxiomC{$r_1 : \CR \> P \> x \> y$}
\AxiomC{$r_2 : \CR \> P \> y \> z$}
\BinaryInfC{$\CRtrans \> r_1 \> r_2 : \CR \> P \> x \> z$}
\end{bprooftree}
\end{center}
\begin{center}
\begin{bprooftree}
\AxiomC{$r : \CR \> P_1 \> x \> y$}
\UnaryInfC{$\CRinl \> r : \CR \> (\sumP{P_1}{P_2}) \> (\inl \> x) \> (\inl \> y)$}
\end{bprooftree}
\begin{bprooftree}
\AxiomC{$r : \CR \> P_2 \> x \> y$}
\UnaryInfC{$\CRinr \> r : \CR \> (\sumP{P_1}{P_2}) \> (\inr \> x) \> (\inr \> y)$}
\end{bprooftree}
\end{center}
\begin{center}
\begin{bprooftree}
\AxiomC{$r_1 : \CR \> P_1 \> x_1 \> x_2$}
\AxiomC{$r_2 : \CR \> P_2 \> y_1 \> y_2$}
\BinaryInfC{$\CRpair \> r_1 \> r_2 : \CR \> (\prodP{P_1}{P_2}) \> (x_1 , x_2) \> (y_1 , y_2)$}
\end{bprooftree}
\end{center}
\begin{center}
\begin{bprooftree}
\AxiomC{$j : \pathI{\sig}$}
\AxiomC{$x : \semP{\pathA{\sig} \> j} \> \CRC$}
\BinaryInfC{$\CRpath \> j \> x : \CR \> (\semE{\pathlh{\sig} \> j} \> \CRC \> x) \> (\semE{\pathrh{\sig} \> j} \> \CRC \> x)$}
\end{bprooftree}
\begin{bprooftree}
\AxiomC{$r : \CR \> (\pointc{\sig}) \> x \> y$}
\UnaryInfC{$\CRcong \> r : \CR \> \I \> (f \> x) \> (f \> y)$}
\end{bprooftree}
\end{center}

Now we define a relation on $\CRC$ by $\CRR \> x \> y = \trunct{\CR \> \I \> x \> y}$.
Because of the truncation, it takes value in proposition.
Note that $\CRR$ is an equivalence relation, because of the constructor $\CRrefl$, $\CRsym$, and $\CRtrans$.
In addition, all equations in $\sig$ are identified in $\CRR$, because of the constructor $\CRpath$.
Lastly, with the constructors $\CRinl$, $\CRinr$, and $\CRcong$, we can show that $\CRR$ preserves $f$.
Hence, we obtain a setoid $I \Def (\CR, \CRC)$ together with a setoid morphism $\varphi : \semPT{P} \> I \rightarrow I$.
To show that this is the initial setoid, we first need a property of $\CRR$, which allows us to make maps from $I$ to algebras.

\begin{lemma}
\label{lem:mapcongrel}
If $g$ is a prealgebra morphism from $\CR$ to some algebra $X$, then $g$ preserves $\CRR$.
\end{lemma}

Now we show that $I$ is an initial object in $\Alg{\sig}$.
To show that we always have a homomorphism from $I$ to an algebra $X$, we use that $\CR$ is the initial prealgebra and lemma \ref{lem:mapcongrel}.
Furthermore, note that two setoid morphisms are equal if their underlying carriers are equal and two algebra morphisms are also equal if their underlying carrier is equal.
Hence, the uniqueness also follows from initiality of $\CR$.
All in all, we get

\begin{theorem}
\label{thm:initialsetoid}
The category $\Algst{\sig}$ has an initial object.
\end{theorem}

From Theorems \ref{thm:algadj} and \ref{thm:initialsetoid} we conclude

\begin{corollary}
\label{thm:hit-exist}
For each signature $\sig$, we can construct a HIT on $\sig$.
\end{corollary}

@@

\begin{corollary}
\label{cor:hit_unique}
If $H_1$ and $H_2$ are HITs for $\sig$, then the underlying $\sig$-algebras of $H_1$ and $H_2$ are equal.
\end{corollary}

@@

\begin{corollary}
\label{cor:hit_path}
Let $\sig$ be a signature.
Then for each $x, y : \CR$ we have an equivalence $\CRR \> x \> y \weq \classof \> x = \classof \> y$.
\end{corollary}

\section{Conclusion and Further Work}
\label{sec:conclusion}
We showed how to construct finitary set-truncated higher inductive types using the quotient and propositional truncation in Theorem \ref{thm:hit-exist}.
Since truncations and the quotient can be constructed from nonrecursive HITs, this proves that all finitary set-truncated HITs can be constructed from nonrecursive ones.
Besides that, we took advantage of this construction to show uniqueness of HITs in Corollary \ref{cor:hit_unique} and to characterize their path space in Corollary \ref{cor:hit_path}.

For a construction of all finitary HITs, higher path constructors need to be allowed and the HITs need to be interpreted as higher categories instead of sets.
A first step towards that goal, would be adapting this construction to bicategory theory \cite{ahrens2019bicategories,capriotti2017univalent} and constructing higher inductive 1-types using Dybjer's and Moeneclaey's interpretation of such HITs in groupoids \cite{DybjerM18}.
Doing so, would require interpreting the path constructor terms and the path computation rule becomes relevant.
Proving the induction principle from initiality is more difficult then, because equality becomes proof relevant.
In this setting, Corollary \ref{cor:hit_path} is more interesting, because the computation rules of path constructors become nontrivial.

@@

	\bibitem{Doorn16}
	Floris van Doorn, \emph{{Constructing the Propositional Truncation using
			Non-Recursive HITs}}, Proceedings of the 5th {ACM} {SIGPLAN} Conference on
	Certified Programs and Proofs, Saint Petersburg, FL, USA, January 20-22,
	2016, 2016, pp.~122--129.	
\end{thebibliography}
\end{document}
